---
title: "R Markdown Tutorial"
author: "Eric Karsten"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: TRUE
    fig_height: 4
    fig_width: 6
---
Now let's say that we want to have a nice PDF readout of some of our tables and
analyses in R. The tool we will use is Rmarkdown. This allows you to generate a
PDF document with output from R code with very little effort. There are lots of
options for how you run your code chunks so that they turn out nicely.

# Including or not including code chunks

Our first code chunk brings in our dataset, so it's not useful to include that 
in the markdown (and you don't see it below in the PDF output).
```{r importing data, include= F}

# This snippet of code is a little loop that makes my code work on your computer
root <- getwd()
while(basename(root) != "Oeconomica-Health-Cohort") {
  root <- dirname(root)
}

# This line runs the script in your data.R file so that each person can have
# their data in a different place because everyone's file structure will be 
# a little differnt
source(file.path(root, "data.R"))


# These are some fantastic packages that I always load in
# You will need to run "install.packages()" with the name of each package
# For example, to install dplyr, run "install.packages("dplyr")"
library(dplyr)
library(ggplot2)
library(knitr)
library(stargazer)

# This is the step where I actually import the data
df <- read.csv(file.path(ddir, "nhis_00001.csv.gz"))


```

Now the below code ins included in our PDF, but I have opted not to have to see any
of the warning messages that come with viewing it.

```{r data cleaning, warning=F}
sex_codebook <-
  tibble(SEX = c(1,2),
         sex_clean = c("Male","Female"))

educ_codebook <- 
  tibble(EDUC = c(0:22, 97:99),
         educ_clean = c(rep("No Degree", 14),
                        rep("HS Diploma",2),
                        rep("Some College", 3),
                        rep("College Degree", 4),
                        rep(NA, 3)
                        )
         )

slim_df <-
  df %>%
  select(AGE, SEX, EDUC, HEALTH, HEIGHT, WEIGHT) %>%
  sample_frac(.1, replace = F) %>%
  mutate(BMI = WEIGHT/HEIGHT) %>%
  left_join(sex_codebook, by = "SEX") %>%
  left_join(educ_codebook, by = "EDUC")
```


# Including tables

Let's say I now want to tell the story of how men and women get less healthy
as they get older. The first thing I might want to show would be a table with
the average health ratings of men and women compared with how old they are. I use
the kable command to include a nice looking table in the markdown.

```{r including tables}

age_health_gender <-
  slim_df %>%
  group_by(AGE, sex_clean) %>%
  summarise(avg = mean(HEALTH)) %>%
  spread(key = sex_clean, value = avg)

kable(age_health_gender, caption = "Average health as respondents age by gender")

age_health_gender

```

This table is so long it is useless, so let's represent this information differently.

# Including ggpplots

Now I want to further convince my readers by showing the difference in the distributions
of our data for health rating by age and gender. To do this I will create a categorical
age variable.


```{r including ggplots}

slim_df %>%
  group_by(AGE, sex_clean) %>%
  summarise(avg = mean(HEALTH)) %>%
  ggplot(aes(x = AGE, y = avg, color = sex_clean)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position = 'bottom') +
  labs(title = "Average health rating by age and sex",
       x = "Age of Respondent",
       color = "Reported sex",
       y = "Average Health Rating (1 is good)")

slim_df %>%
  mutate(age = case_when(AGE < 20 ~ "20 and below",
                         AGE < 40 ~ "20 to 40",
                         AGE < 60 ~ "40 to 60",
                         AGE < 90 ~ "60 to 90")) %>%
  ggplot(aes(x = HEALTH, fill = sex_clean)) +
  geom_bar(position = 'fill')+
  facet_grid(. ~ age) +
  theme(legend.position = 'bottom') +
  labs(title = "Distributions of health rating by age and sex",
       x = "Health Rating",
       fill = "Reported sex",
       y = "Gender ratio for reporting this rating")


```

# Including regression output

Now the below code chunk has a few options. The `results='asis'` option is there
to make the output of stargazer render nicely. The `echo = F` prevents the source
code form being inlcuded, and the `warning = F` prevents any warnings form being included
in the PDF.

```{r including models, results='asis', echo= F, warning= F}


model_1 <- lm(HEALTH ~ sex_clean + AGE, data = slim_df)

model_2 <- lm(HEALTH ~ sex_clean * AGE, data = slim_df)

stargazer(model_1, model_2)

```
