---
title: "The Case for data.table"
author: "Eric Karsten"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  beamer_presentation: 
    theme: Singapore
---

# The Options

## Base R

Pros:

  - It's the first thing you learn
  
Cons:

  - Code is unreadable
  - Iterative processes require for loops (unclear) or
  - calling the `apply` function (slow)


## dplyr and the tidyverse

Pros:

  - The syntax is very readable
  - Major speed improvements over base R
  - More complex aggregations possible
  
Cons:

  - You will eventually come across aggregations you can't do
  - It's a little slow when you start doing big stuff
  - You can't poke around the backend code easily (it's in C++)


## data.table

Pros:

  - Almost every aggregation you can think of is possible
  - Some speed improvements over dplyr (task-dependent)
  - Significanlty more efficient use of RAM than dplyr or base R
  
Cons:
  
  - Sytax is hard to read
  - You have to know what you are doing to get the most efficiency out of it
  (this presentation will help with that!)
  - You can't poke around the backend code easily (it's in C++)

# Simple Example


## Our Data

This is a super simple example so that you can see all the parts.

```{r data}
# you may need to `install.packages("data.table")`

library(data.table)

df <-
  data.table(
    Animal = c("Dog", "Cat", "Dog",
               "Raven", "Cat"), 
    Weight = c(100, 40, 80, 16, 50),
    Height = c(23, 18, 40, 3, 16),
    Family = c("Mammal", "Mammal", "Mammal",
               "Bird", "Mammal"))
```


## The Syntax

There are three parts to a data.table command:

 1. The rows to be returned
 2. The columns to be returned
 3. The grouping by which to perform the operations
 
They are separated by commas, but commas need only be included when a subsequent
field is occupied.

## Some nice easy subsetting

Let's just select the dogs.

```{r subsetting a}

df[Animal == "Dog"]

```

Let's return the height and weight of the mammals that weigh more than 35 lbs

```{r subsetting b}

df[Weight > 35, .(Animal, Height, Weight)]

```


## Using `:=`

Often times, we want to add new calculated columns to a dataframe. The correct
way to do this in data.table is with a `:=`. Because this operation works by
updating the previous object, we need to make
a copy of it in order to avoid modifying our original data frame.

The following are equivalent for computing BMI (weight/height) \footnote{BMI
should technically be weight in Kg divided by height in m, but I'm not going to
worry about the conversions}

```{r fast-assign}
df2 <- copy(df)[, BMI := Weight/Height]
df2 <- copy(df)[, `:=`(BMI = Weight/Height)]
```

```{r, echo=FALSE}
print(df2)
```

---

We could also have done this without efficient updating. This is much slower
on big datasets and should not be done.

```{r slow-update}

df[, .(Animal, Weight, Height, Family,
       BMI = Weight/Height)]

```
---

Selective updating is the reason why `:=` is so useful. Let's say we wanted to
change the hieght of all ravens to be 5. We don't want to rewrite the whole table,
so we simply call

```{r selective-update}

df[Animal == "Raven", Height := 5]

```

```{r, echo=FALSE}
print(df)
```


## Grouped operations

Let's say we want the average weight by animal. Here are two ways to do it with
different results.

```{r grouping}

df[, .(mean_wt = mean(Weight)), by = .(Animal)]
df2 <- copy(df)[, mean_wt := mean(Weight), by = Animal]

```

```{r, echo=FALSE}
print(df2)
```

## Chaining multiple operations in sequence

Oftentimes, you want to do several operations in sequence. Naturally, there is
syntax that reflects this workflow. Here we are computing BMI and then making a
table of average BMI by Animal.

```{r chaining}

df[, `:=`(BMI = Weight/Height)][
  , .(avg_BMI = mean(BMI)), by = Animal]

```

It should be noted that this example could be done in one line, so chaining was
not necessary.

## Using `.SD` tricks

There are times when we want to perform more complex operations on our grouped
data than applying a simple built-in function. One option is to write a function
of our own, but this isn't always the clearest. The better option is to manipulate
the grouped data.table using standard data.table operations!

The way we do this is by calling the `.SD` object (which is just a data.table of
our selected data) and performing operations on it.

EX: We want to add a column for the height of the heaviest individual within each
Animal species.

To do this, we need to first compute within each animal table what the biggest
weight is, then, we need to return the height of the animal with that weight
and assign it to the BigHeight variable for that animal group.

---

```{r fancy-aggregation}

df[, BigHeight := .SD[, .(BiggestWeight = max(Weight),
                           Weight,
                           Height)][
                       Weight == BiggestWeight, Height]
   , by = Animal]

```

```{r, echo=FALSE}
print(df)
```

I challenge you to do this in either dplyr or base R in a cleaner or faster way!

# Speed Tricks

## Why is it so fast?

Both data.table and dplyr are executed in C++. This allows the developers to
use some really efficient indexing to perform their operations.

The advantage data.table has over dplyr in the speed department is that it is
able to update the data in RAM, it doesn't need to constantly be rewriting it.
This saves a ton of time if you are doing thousands of small operations in a row.
It also allows the program to run without eating up all your RAM!

## How to maximize speed

Always use `:=` when possible. You never want to rewrite your data.table
if you don't have to.

To delete a column quickly, just use `:= NULL` to get rid of it.

If you are doing several things on the same group, don't chain the operations
together, instead do them all at once.

Merging isn't covered in this talk, but data.table can do really fast merging.

# Proof

## The task

We are going to use the microbenchmark library to find the average BMI of our
animals, after adding 7 lbs to all the rabbits but to slow it down,
we are going to randomly generate a ton of data.

```{r, message=F, warning=F}

library(microbenchmark)
library(dplyr)
n <- 1e5
dt_test <-
  data.table(Animal = sample(c("Rabbit", "Chicken",
                               "Zebra", "Snake",
                               "Giraffe", "Eagle"),
                             n, replace = T),
             Height = runif(n, 3, 7),
             Weight = runif(n, 33, 47))
df_test <- as.data.frame(dt_test)
tb_test <- as_tibble(dt_test)

```

## The Benchmark Setup

```{r benchmark}

data_table_BMI <- function(df) {
  df[Animal == "Rabbit", Weight := Weight + 7][
    , .(avg_BMI= mean(Weight/Height)), by = Animal]}
dplyr_BMI <- function(df) {
  df %>%
    mutate(Weight = if_else(Animal == "Rabbit",
                            Weight + 7,Weight)) %>%
    group_by(Animal) %>%
    summarise(avg_BMI = mean(Weight/Height))}
base_R_BMI <- function(df) {
  df$Weight = ifelse(df$Animal == "Rabbit",
                     df$Weight + 7,df$Weight)
  aggregate(df$Weight/df$Height,
            by=list(Animal=df$Animal),
            FUN=mean)}

```

## Benchmark Results

```{r print me}
res <-
  microbenchmark(
    data_table_BMI(dt_test),
    dplyr_BMI(tb_test),
    base_R_BMI(df_test))

```

```{r benchmark results, echo=F, message=FALSE, warning=FALSE}

codebook <-
  tibble(expr = c("base_R_BMI(df_test)", "data_table_BMI(dt_test)","dplyr_BMI(tb_test)" ),
         Method = c("Base R", "data.table", "dplyr"))

res %>%
  left_join(codebook, by = "expr") %>%
  mutate(time = time/1000000) %>%
  group_by(Method) %>%
  summarise(Mean = round(mean(time), digits = 2),
            Median = round(median(time), digits = 2),
            SD = round(sd(time), digits = 2))

```

# Concluding Thoughts

## A Handy Guide

- If you data is less than 100 rows, skip the computer and do it with pencil and paper,
 it really doesn't even matter
- Less than 1000 and everything will feel fast
- Less than 1,000,000 and dplyr should be alright
- Less than 10,000,000 and data.table will work on your laptop
- Otherwise, it's time to go shopping for server space

## Thank You

A much more comprehensive data.table guide is provided here:
https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html


